---
title: "Gov 50 Final Project"
author: "Sarah Bradley"
description: "Investigating the relationship between ethnolinguistic fractionalism and constitutional change"
output:
  distill::distill_article:
    self_contained: false
---



## Introduction

In this project I will explore the relationship between ethnolinguistic fractionalism and constitutional change. I have chosen to explore ethnolinguistic fractionalism as an independent variable instead of ethnic separatism (although I suspect there would be an even stronger positive correlation there) as it is a less studied area of research. The dependent variable I will be researching is major constitutional change, specifically, the number of basic alterations to a state's constitution. I hypothesize that there will be a positive correlation between the two. The null hypothesis would be that there is no relationship.

## The Data

I am using a correlational research data 

```{r}
file_path <- "~/Downloads/violence.csv"

data <- read.csv(file_path)


#linear regression

model <- lm(constchg60 ~ elf60, data = data)

summary(model)

#tibble

library(tibble)
library(broom)

model <- lm(constchg60 ~ elf60, data = data)

regression_results <- tidy(model)

print(regression_results)


```
```{r}
#investigating confoudning variables


model <- lm(constchg60 ~ govtcris60 + revols60 + elf60, data = data)

summary(model)


```

```{r}

# Load ggplot2 package
library(ggplot2)

# Remove rows with N/A values for constchg60
data_no_na_constchg60 <- data[complete.cases(data$constchg60), ]

# Create a bar chart with custom labels
ggplot(data_no_na_constchg60, aes(x = factor(constchg60), fill = factor(elf60 >= 0.5))) +
  geom_bar(position = "dodge", alpha = 0.7) +
  labs(title = "Ethnolinguistic Fractionalism and Constitutional Change",
       x = "Constitutional Change",
       y = "Number of Countries") +
  scale_fill_manual(values = c("TRUE" = "blue", "FALSE" = "red"), 
                    name = "Ethnolinguistic Fractionalization > 0.5?") +
  scale_x_discrete(labels = c("0" = "No Change", "1" = "Change")) +
  theme_minimal()

print(geom_bar)
```
```{r}
# Load ggplot2 package
library(ggplot2)

# Remove rows with N/A values for constchg60
data_no_na_constchg60 <- data[complete.cases(data$constchg60), ]

# Create a box plot
ggplot(data_no_na_constchg60, aes(x = factor(constchg60), y = elf60)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7) +
  labs(title = "Ethnolinguistic Fractionalism and Constitutional Change",
       x = "Constitutional Change",
       y = "Ethnolinguistic Fractionalization") +
  scale_x_discrete(labels = c("0" = "No Change", "1" = "Change")) +
  theme_minimal()

print(ggplot)

```



```{r}
## logistical regression

logistic_model <- glm(constchg60 ~ elf60, data = data, family = "binomial")

summary(logistic_model)

```

## Results

## Conclusion
